\pkg{hesim}: Health Economic Simulation Modeling and Decision Analysis

### 2021-03-07 manuscript.tex
Add files needed for arXiv submission

\documentclass[article, nojss]{jss}

@@

<<include=FALSE>>=
library(knitr)
knitr::render_sweave()
opts_chunk$set(
  engine="R", tidy = FALSE, eval = TRUE, cache = TRUE,
  fig.height = 4, fig.width = 6
)
@

<<preliminaries, echo=FALSE, results="hide">>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
set.seed(102)
@

@@

  Jeroen P Jansen\\
  University of California, San Francisco\\
  San Francisco\\
  E-mail: \email{jeroen.jansen@ucsf.edu}\\
}

\begin{document}

@@

\begin{table} [h]
\footnotesize
\begin{tabular*}{\textwidth}{l l l l}
\hline
Economic model & Statistical model & Parameter object & Model object\\
\hline
\code{CohortDtstm} & Custom & \code{tparams_transprobs} & \code{msm::msm} \\
 & Multinomial logistic regressions & \code{params_mlogit} & \code{multinom_list}\\
\code{IndivCtstm} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
& Multi-state model (transition-specific) & \code{params_surv_list} & \code{flexsurvreg_list}\\
\code{Psm} & Independent survival models & \code{param_surv_list} & \code{flexsurvreg_list}\\
\hline
\end{tabular*}
\caption{Parameterization of disease models.}\label{tbl:parameterize-disease-model}
\end{table}

@@

\begin{table} [h]
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Statistical model & Parameter object & Model object\\
\hline
Predicted means & \code{tparams_mean} & \code{stateval_tbl}\\
Linear model & \code{params_lm} & \code{stats::lm}\\
\hline
\end{tabular*}
\caption{Parameterization of state value models.}\label{tbl:parameterize-stateval-model}
\end{table}

@@

\begin{table} [h]
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease model & Utility model & Cost model(s)\\
\hline
\code{CohortDtstm} & \code{CohortDtstmTrans} & \code{StateVals} & \code{StateVals}\\
\code{IndivCtstm} & \code{IndivCtstmTrans} & \code{StateVals} & \code{StateVals}\\
\code{Psm} & \code{PsmCurves} & \code{StateVals} & \code{StateVals}\\
\hline
\end{tabular*}
\caption{Submodels comprising an economic model.}\label{tbl:submodels}
\end{table}

@@

\begin{table} [h]
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease progression  & QALYs & Costs \\
\hline
\code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\hline
\end{tabular*}
\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
\end{table}

@@

<<tmat>>=
tmat <- rbind(
  c(NA, 1, 2),
  c(NA, NA, 3),
  c(NA, NA, NA)
)
colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
print(tmat)
@

@@

<<hesim_data>>=
library("hesim")
library("data.table")
n_patients <- 1000
patients <- data.table(
  patient_id = 1:n_patients,
  age = rnorm(n_patients, mean = 45, sd = 7),
  female = rbinom(n_patients, size = 1, prob = .51)
)

states <- data.table(
  state_id = c(1, 2),
  state_name = c("Stable", "Progression") 
)

strategies <- data.table(
  strategy_id = 1:3,
  strategy_name = c("SOC", "New 1", "New 2"),
  soc = c(1, 0, 0),
  new1 = c(0, 1, 0),
  new2 = c(0, 0, 1)
)

hesim_dat <- hesim_data(
  strategies = strategies,
  patients = patients,
  states = states
)
print(hesim_dat)
@

@@

<<labels_indiv>>=
labs_indiv <- get_labels(hesim_dat)
print(labs_indiv)
@

@@

<<ictstm_data>>=
onc3[patient_id %in% c(1, 2)]
@

@@

<<ictstm_fits, warning = FALSE, message = FALSE>>=
library("flexsurv")
n_trans <- max(tmat, na.rm = TRUE) 
wei_fits <- vector(length = n_trans, mode = "list")
f <- as.formula(Surv(time, status) ~ factor(strategy_name) + female + age)
for (i in 1:length(wei_fits)){
  if (i == 3) f <- update(f, .~.-factor(strategy_name)) 
  wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
                               subset = (transition_id == i),
                               dist = "weibull")
}
wei_fits <- flexsurvreg_list(wei_fits)
@

@@

<<ictstm_utility>>=
utility_tbl <- stateval_tbl(
  data.table(state_id = states$state_id,
             mean = c(.8, .6),
             se = c(0.02, .05)
  ),
  dist = "beta")
print(utility_tbl)
@

@@

<<ictstm_medcosts>>=
medcost_tbl <- stateval_tbl(
  data.table(state_id = states$state_id,
             mean = c(2000, 9500),
             se = c(2000, 9500)
  ),
  dist = "gamma")
print(medcost_tbl)
@

@@

<<ictstm_drugcost_setup, echo=FALSE, results="hide">>=
n_times <- 2
n_states <- nrow(states)
n_strategies <- nrow(strategies)
n_i <- n_states * n_times
drugcost_dt <-  data.table(
    strategy_id = rep(strategies$strategy_id, each = n_i),
    state_id = rep(rep(states$state_id, each = n_times), n_strategies),
    time_start = rep(c(0, 3/12), n_states * n_strategies),
    est = c(2000, 2000, 1500, 1200,
           12000, 12000, 1500, 1200,
           15000, 15000, 1500, 1200)
) 
@

<<ictstm_drugcost>>=
drugcost_tbl <- stateval_tbl(
  drugcost_dt,
  dist = "fixed")
print(drugcost_tbl)
@

@@

<<ictstm_n_samples>>=
n_samples <- 1000
@

@@

<<ictstm_transmod_data>>=
transmod_data <- expand(hesim_dat,
                        by = c("strategies", "patients"))
head(transmod_data)
@

@@

<<ictstm_transmod>>=
transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
                                   trans_mat = tmat, n = n_samples,
                                   uncertainty = "normal",
                                   clock = "reset",
                                   start_age = patients$age)
@

@@

<<ictstm_stvalmods>>=
utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
                               hesim_data = hesim_dat)
drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
                                time_reset = TRUE, hesim_data = hesim_dat)
medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
                               hesim_data = hesim_dat)
costmods <- list(Drug = drugcostmod,
                 Medical = medcostmod)
@

@@

<<ictstm_econmod>>=
ictstm <- IndivCtstm$new(trans_model = transmod,
                         utility_model = utilitymod,
                         cost_models = costmods)
@

@@

<<ictstm_sim_disease>>=
ictstm$sim_disease(max_age = 100)
head(ictstm$disprog_)
@

@@

<<ictstm_sim_stateprobs>>=
ictstm$sim_stateprobs(t = seq(0, 30 , 1/12))
head(ictstm$stateprobs_)
@

@@

<<ictstm_plot_stateprobs, fig.show = "hide">>=
library("ggplot2")
autoplot(ictstm$stateprobs_, labels = labs_indiv,
         ci = FALSE) 
@

@@

<<ictstm_sim_qalys>>=
ictstm$sim_qalys(dr = c(0,.03))
head(ictstm$qalys_)
@

<<ictstm_sim_costs>>=
ictstm$sim_costs(dr = .03)
head(ictstm$costs_)
@

@@

<<ictstm_summarize>>=
library("magrittr")
ce_sim_ictstm <- ictstm$summarize()
summary(ce_sim_ictstm, labels = labs_indiv) %>%
  format()
@

@@

<<psm_hesim_data>>=
hesim_dat$patients <- data.table(
  patient_id = 1:4,
  grp_id = c(1, 1, 2, 2),
  patient_wt = rep(1/4, 4),
  age = c(55, 65, 55, 65),
  female = c(1, 1, 0, 0),
  grp_name = rep(c("Female", "Male"), 
                 each = 2)
)
@

@@

<<labels_cohort>>=
labs_cohort <- get_labels(hesim_dat)
@

@@

<<psm_data>>=
onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
                                               "strategy_name"))
onc_pfs_os[patient_id %in% c(1, 2)]
@

@@

<<psm_fits_soc>>=
onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
fit_pfs_soc_wei <- flexsurvreg(
  Surv(pfs_time, pfs_status) ~ female,
  data = onc_pfs_os_soc,
  dist = "weibull")

fit_os_soc_wei <- flexsurvreg(
  Surv(os_time, os_status) ~  female,
  data = onc_pfs_os_soc,
  dist = "weibull")

psmfit_soc_wei <- partsurvfit(
  flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
  data = onc_pfs_os_soc
)
@

@@

<<psm_fits, include = FALSE>>=
# Create dummies
onc_pfs_os[, new1 := ifelse(strategy_name == "New 1", 1, 0)]
onc_pfs_os[, new2 := ifelse(strategy_name == "New 2", 1, 0)]

# Fit models
fit_pfs_wei <- flexsurv::flexsurvreg(
  Surv(pfs_time, pfs_status) ~ female + new1 + new2,
  data = onc_pfs_os,
  dist = "weibull")

fit_os_wei <- flexsurvreg(
  Surv(os_time, os_status) ~ female + new1 + new2,
  data = onc_pfs_os,
  dist = "weibull")

# Extract relative treatment effects
params_pfs_wei <- create_params(fit_pfs_wei, n = n_samples)
params_os_wei <- create_params(fit_os_wei, n = n_samples)
coef_pfs_wei <- params_pfs_wei$coefs$scale[, c("new1", "new2")]
coef_os_wei <- params_os_wei$coefs$scale[, c("new1", "new2")]
@

<<psm_coefs>>=
head(coef_pfs_wei)
@

@@

<<psm_drugcost>>=
drugcost_tbl <- stateval_tbl(
  drugcost_dt[time_start ==  0][, time_start := NULL],
  dist = "fixed")
print(drugcost_tbl)
@

@@

<<psm_survmods_data>>=
survmods_data <- expand(hesim_dat, by = c("strategies", "patients"))
head(survmods_data)
@

@@

<<psm_survmods_params, warning = FALSE>>=
survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
                                 uncertainty = "bootstrap", 
                                 max_errors = 5, silent = TRUE)
survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
                                         coef_pfs_wei)
survmods_params$os$coefs$scale <- cbind(survmods_params$os$coefs$scale,
                                        coef_os_wei)
survmods_params$pfs$coefs$scale[1:2, ]
@

@@

<<psm_survmods>>=
survmods_data[, ("(Intercept)") := 1]
survmods <- create_PsmCurves(survmods_params, 
                             input_data = survmods_data)
@

@@

<<psm_stvalmods>>=
utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
                               hesim_data = hesim_dat)
drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples, 
                                hesim_data = hesim_dat)
medcostmod <- create_StateVals(medcost_tbl, n = n_samples, 
                               hesim_data = hesim_dat)
costmods <- list(Drug = drugcostmod, Medical = medcostmod)
@

@@

<<psm_econmod>>=
psm <- Psm$new(survival_models = survmods,
               utility_model = utilitymod,
               cost_models = costmods)
@

@@

<<psm_sim_survival>>=
times <- seq(0, 30, by = .1)
psm$sim_survival(t = times)
head(psm$survival_)
@

<<psm_plot_survival, fig.show = "hide">>=
autoplot(psm$survival_, 
         labels = c(labs_cohort,
                    list(curve = c("PFS" = 1, "OS" = 2))
                    ),
         ci = TRUE, ci_style = "ribbon")
@

@@

<<psm_sim_stateprobs>>=
psm$sim_stateprobs()
psm$stateprobs_[sample == 1 & patient_id == 1 & state_id == 2 & t == 12]
@

@@

<<psm_costs_qalys>>=
psm$sim_qalys(dr = .03) 
psm$sim_costs(dr = .03)
@

<<psm_summary>>=
ce_sim_psm <- psm$summarize(by_grp = TRUE)
summary(ce_sim_psm, labels = labs_cohort) %>%
  format()
@

>>>>

\documentclass[article, nojss]{jss}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\usepackage{Sweave}

@@

@@

  Jeroen P Jansen\\
  University of California, San Francisco\\
  San Francisco\\
  E-mail: \email{jeroen.jansen@ucsf.edu}\\
}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

\begin{table} [h]
\caption{Parameterization of disease models.}\label{tbl:parameterize-disease-model}
\footnotesize
\begin{tabular*}{\textwidth}{l l l l}
\hline
Economic model & Statistical model & Parameter object & Model object\\
\hline
\code{CohortDtstm} & Custom & \code{tparams_transprobs} & \code{msm::msm} \\
 & Multinomial logistic regressions & \code{params_mlogit} & \code{multinom_list}\\
\code{IndivCtstm} & Multi-state model (joint likelihood) & \code{params_surv} & \code{flexsurv::flexsurvreg}\\
& Multi-state model (transition-specific) & \code{params_surv_list} & \code{flexsurvreg_list}\\
\code{Psm} & Independent survival models & \code{param_surv_list} & \code{flexsurvreg_list}\\
\hline
\end{tabular*}
\end{table}

@@

\begin{table} [h]
\caption{Parameterization of state value models.}\label{tbl:parameterize-stateval-model}
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Statistical model & Parameter object & Model object\\
\hline
Predicted means & \code{tparams_mean} & \code{stateval_tbl}\\
Linear model & \code{params_lm} & \code{stats::lm}\\
\hline
\end{tabular*}
\end{table}

@@

\begin{table} [h]
\caption{Submodels comprising an economic model.}\label{tbl:submodels}
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease model & Utility model & Cost model(s)\\
\hline
\code{CohortDtstm} & \code{CohortDtstmTrans} & \code{StateVals} & \code{StateVals}\\
\code{IndivCtstm} & \code{IndivCtstmTrans} & \code{StateVals} & \code{StateVals}\\
\code{Psm} & \code{PsmCurves} & \code{StateVals} & \code{StateVals}\\
\hline
\end{tabular*}
\end{table}

@@

\begin{table} [h]
\caption{Methods for simulating outcomes from economic models.}\label{tbl:simulate-outcomes}
\footnotesize
\begin{tabular*}{\textwidth}{@{\extracolsep{\fill}}l l l l}
\hline
Economic model & Disease progression  & QALYs & Costs \\
\hline
\code{CohortDtstm} & \code{sim_stateprobs()} & \code{sim_qalys()} & \code{sim_costs()}\\
\code{IndivCtstm} & \code{sim_disease()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\code{Psm} & \code{sim_survival()}, \code{sim_stateprobs()}  & \code{sim_qalys()} & \code{sim_costs()}\\
\hline
\end{tabular*}
\end{table}

@@

\begin{Schunk}
\begin{Sinput}
R> tmat <- rbind(
+    c(NA, 1, 2),
+    c(NA, NA, 3),
+    c(NA, NA, NA)
+  )
R> colnames(tmat) <- rownames(tmat) <- c("Stable", "Progression", "Dead")
R> print(tmat)
\end{Sinput}
\begin{Soutput}
            Stable Progression Dead
Stable          NA           1    2
Progression     NA          NA    3
Dead            NA          NA   NA
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> library("hesim")
> library("data.table")
> n_patients <- 1000
> patients <- data.table(
+   patient_id = 1:n_patients,
+   age = rnorm(n_patients, mean = 45, sd = 7),
+   female = rbinom(n_patients, size = 1, prob = .51)
+ )
> 
> states <- data.table(
+   state_id = c(1, 2),
+   state_name = c("Stable", "Progression") 
+ )
> 
> strategies <- data.table(
+   strategy_id = 1:3,
+   strategy_name = c("SOC", "New 1", "New 2"),
+   soc = c(1, 0, 0),
+   new1 = c(0, 1, 0),
+   new2 = c(0, 0, 1)
+ )
> 
> hesim_dat <- hesim_data(
+   strategies = strategies,
+   patients = patients,
+   states = states
+ )
> print(hesim_dat)
\end{Sinput}
\begin{Soutput}
$strategies
   strategy_id strategy_name soc new1 new2
1:           1           SOC   1    0    0
2:           2         New 1   0    1    0
3:           3         New 2   0    0    1

$patients
      patient_id      age female
   1:          1 46.26366      1
   2:          2 50.49314      1
   3:          3 35.52785      1
   4:          4 58.88309      0
   5:          5 53.66930      0
  ---                           
 996:        996 54.60555      0
 997:        997 50.10654      0
 998:        998 45.02035      0
 999:        999 45.60573      1
1000:       1000 39.45426      0

$states
   state_id  state_name
1:        1      Stable
2:        2 Progression

attr(,"class")
[1] "hesim_data"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> labs_indiv <- get_labels(hesim_dat)
R> print(labs_indiv)
\end{Sinput}
\begin{Soutput}
$strategy_id
  SOC New 1 New 2 
    1     2     3

$state_id
     Stable Progression       Death 
          1           2           3 
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> onc3[patient_id %in% c(1, 2)]
\end{Sinput}
\begin{Soutput}
          from          to strategy_name female      age patient_id
1:      Stable Progression         New 2      0 59.85813          1
2:      Stable       Death         New 2      0 59.85813          1
3: Progression       Death         New 2      0 59.85813          1
4:      Stable Progression         New 2      0 62.57282          2
5:      Stable       Death         New 2      0 62.57282          2
   time_start time_stop status transition_id strategy_id      time
1:   0.000000  2.420226      1             1           3  2.420226
2:   0.000000  2.420226      0             2           3  2.420226
3:   2.420226 14.620258      1             3           3 12.200032
4:   0.000000  7.497464      0             1           3  7.497464
5:   0.000000  7.497464      0             2           3  7.497464
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> library("flexsurv")
R> n_trans <- max(tmat, na.rm = TRUE) 
R> wei_fits <- vector(length = n_trans, mode = "list")
R> f <- as.formula(Surv(time, status) ~ factor(strategy_name) + female + age)
R> for (i in 1:length(wei_fits)){
+    if (i == 3) f <- update(f, .~.-factor(strategy_name)) 
+    wei_fits[[i]] <- flexsurvreg(f, data = onc3, 
+                                 subset = (transition_id == i),
+                                 dist = "weibull")
+  }
R> wei_fits <- flexsurvreg_list(wei_fits)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> utility_tbl <- stateval_tbl(
+    data.table(state_id = states$state_id,
+               mean = c(.8, .6),
+               se = c(0.02, .05)
+    ),
+    dist = "beta")
R> print(utility_tbl)
\end{Sinput}
\begin{Soutput}
   state_id mean   se
1:        1  0.8 0.02
2:        2  0.6 0.05
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> medcost_tbl <- stateval_tbl(
+    data.table(state_id = states$state_id,
+               mean = c(2000, 9500),
+               se = c(2000, 9500)
+    ),
+    dist = "gamma")
R> print(medcost_tbl)
\end{Sinput}
\begin{Soutput}
   state_id mean   se
1:        1 2000 2000
2:        2 9500 9500
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> drugcost_tbl <- stateval_tbl(
+    drugcost_dt,
+    dist = "fixed")
R> print(drugcost_tbl)
\end{Sinput}
\begin{Soutput}
    strategy_id state_id time_id time_start time_stop   est
 1:           1        1       1       0.00      0.25  2000
 2:           1        1       2       0.25       Inf  2000
 3:           1        2       1       0.00      0.25  1500
 4:           1        2       2       0.25       Inf  1200
 5:           2        1       1       0.00      0.25 12000
 6:           2        1       2       0.25       Inf 12000
 7:           2        2       1       0.00      0.25  1500
 8:           2        2       2       0.25       Inf  1200
 9:           3        1       1       0.00      0.25 15000
10:           3        1       2       0.25       Inf 15000
11:           3        2       1       0.00      0.25  1500
12:           3        2       2       0.25       Inf  1200
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> n_samples <- 1000
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> transmod_data <- expand(hesim_dat,
+                          by = c("strategies", "patients"))
R> head(transmod_data)
\end{Sinput}
\begin{Soutput}
   strategy_id patient_id strategy_name soc new1 new2      age female
1:           1          1           SOC   1    0    0 46.26366      1
2:           1          2           SOC   1    0    0 50.49314      1
3:           1          3           SOC   1    0    0 35.52785      1
4:           1          4           SOC   1    0    0 58.88309      0
5:           1          5           SOC   1    0    0 53.66930      0
6:           1          6           SOC   1    0    0 53.40432      1
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> transmod <- create_IndivCtstmTrans(wei_fits, transmod_data,
+                                     trans_mat = tmat, n = n_samples,
+                                     uncertainty = "normal",
+                                     clock = "reset",
+                                     start_age = patients$age)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
+                                hesim_data = hesim_dat)
> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples,
+                                 time_reset = TRUE, hesim_data = hesim_dat)
> medcostmod <- create_StateVals(medcost_tbl, n = n_samples,
+                                hesim_data = hesim_dat)
> costmods <- list(Drug = drugcostmod,
+                  Medical = medcostmod)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> ictstm <- IndivCtstm$new(trans_model = transmod,
+                           utility_model = utilitymod,
+                           cost_models = costmods)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> ictstm$sim_disease(max_age = 100)
R> head(ictstm$disprog_)
\end{Sinput}
\begin{Soutput}
   sample strategy_id patient_id grp_id from to final time_start
1:      1           1          1      1    1  3     1   0.000000
2:      1           1          2      1    1  2     0   0.000000
3:      1           1          2      1    2  3     1   8.058198
4:      1           1          3      1    1  2     0   0.000000
5:      1           1          3      1    2  3     1  15.250551
6:      1           1          4      1    1  3     1   0.000000
   time_stop
1: 11.183651
2:  8.058198
3: 12.794488
4: 15.250551
5: 23.791357
6:  6.047286
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> ictstm$sim_stateprobs(t = seq(0, 30 , 1/12))
R> head(ictstm$stateprobs_)
\end{Sinput}
\begin{Soutput}
   sample strategy_id grp_id state_id          t  prob
1:      1           1      1        1 0.00000000 1.000
2:      1           1      1        1 0.08333333 1.000
3:      1           1      1        1 0.16666667 1.000
4:      1           1      1        1 0.25000000 0.999
5:      1           1      1        1 0.33333333 0.998
6:      1           1      1        1 0.41666667 0.996
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> library("ggplot2")
R> autoplot(ictstm$stateprobs_, labels = labs_indiv,
+           ci = FALSE) 
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> ictstm$sim_qalys(dr = c(0,.03))
R> head(ictstm$qalys_)
\end{Sinput}
\begin{Soutput}
   sample strategy_id grp_id state_id dr    qalys      lys
1:      1           1      1        1  0 4.933188 6.374648
2:      1           1      1        2  0 3.233055 5.255885
3:      1           2      1        1  0 5.604984 7.242740
4:      1           2      1        2  0 2.683823 4.363013
5:      1           3      1        1  0 6.289275 8.126979
6:      1           3      1        2  0 2.915707 4.739981
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> ictstm$sim_costs(dr = .03)
R> head(ictstm$costs_)
\end{Sinput}
\begin{Soutput}
   sample strategy_id grp_id state_id   dr category      costs
1:      1           1      1        1 0.03     Drug  11352.373
2:      1           1      1        2 0.03     Drug   4691.177
3:      1           2      1        1 0.03     Drug  76292.008
4:      1           2      1        2 0.03     Drug   3819.849
5:      1           3      1        1 0.03     Drug 105209.810
6:      1           3      1        2 0.03     Drug   4042.658
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> library("magrittr")
R> ce_sim_ictstm <- ictstm$summarize()
R> summary(ce_sim_ictstm, labels = labs_indiv) %>%
+    format()
\end{Sinput}
\begin{Soutput}
   Discount rate        Outcome                      SOC
1:          0.00          QALYs        7.87 (7.12, 8.60)
2:          0.03          QALYs        6.58 (6.03, 7.14)
3:          0.03    Costs: Drug  15,352 (14,431, 16,234)
4:          0.03 Costs: Medical  46,784 (4,518, 144,256)
5:          0.03   Costs: total 62,136 (19,888, 159,037)
                       New 1                      New 2
1:         8.59 (7.81, 9.40)         9.20 (8.42, 10.06)
2:         7.10 (6.52, 7.69)          7.52 (6.96, 8.11)
3:   78,465 (73,854, 83,200)  105,306 (99,228, 111,686)
4:   46,222 (4,543, 137,303)    46,840 (4,746, 143,093)
5: 124,687 (82,038, 218,484) 152,146 (109,625, 245,279)
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> hesim_dat$patients <- data.table(
+    patient_id = 1:4,
+    grp_id = c(1, 1, 2, 2),
+    patient_wt = rep(1/4, 4),
+    age = c(55, 65, 55, 65),
+    female = c(1, 1, 0, 0),
+    grp_name = rep(c("Female", "Male"), 
+                   each = 2)
+  )
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> labs_cohort <- get_labels(hesim_dat)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> onc_pfs_os <- as_pfs_os(onc3, patient_vars = c("patient_id", "female", "age", 
+                                                 "strategy_name"))
R> onc_pfs_os[patient_id %in% c(1, 2)]
\end{Sinput}
\begin{Soutput}
   patient_id female      age strategy_name pfs_time pfs_status
1:          1      0 59.85813         New 2 2.420226          1
2:          2      0 62.57282         New 2 7.497464          0
   os_status   os_time
1:         1 14.620258
2:         0  7.497464
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> onc_pfs_os_soc <- onc_pfs_os[strategy_name == "SOC"]
> fit_pfs_soc_wei <- flexsurvreg(
+   Surv(pfs_time, pfs_status) ~ female,
+   data = onc_pfs_os_soc,
+   dist = "weibull")
> 
> fit_os_soc_wei <- flexsurvreg(
+   Surv(os_time, os_status) ~  female,
+   data = onc_pfs_os_soc,
+   dist = "weibull")
> 
> psmfit_soc_wei <- partsurvfit(
+   flexsurvreg_list(pfs = fit_pfs_soc_wei, os = fit_os_soc_wei),
+   data = onc_pfs_os_soc
+ )
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> head(coef_pfs_wei)
\end{Sinput}
\begin{Soutput}
          new1      new2
[1,] 0.1814189 0.2654431
[2,] 0.1390351 0.2907570
[3,] 0.1298460 0.2307868
[4,] 0.1729702 0.3143110
[5,] 0.1988663 0.3179721
[6,] 0.1305311 0.2261630
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> drugcost_tbl <- stateval_tbl(
+    drugcost_dt[time_start ==  0][, time_start := NULL],
+    dist = "fixed")
R> print(drugcost_tbl)
\end{Sinput}
\begin{Soutput}
   strategy_id state_id   est
1:           1        1  2000
2:           1        2  1500
3:           2        1 12000
4:           2        2  1500
5:           3        1 15000
6:           3        2  1500
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> survmods_data <- expand(hesim_dat, by = c("strategies", "patients"))
R> head(survmods_data)
\end{Sinput}
\begin{Soutput}
   strategy_id patient_id strategy_name soc new1 new2 grp_id
1:           1          1           SOC   1    0    0      1
2:           1          2           SOC   1    0    0      1
3:           1          3           SOC   1    0    0      2
4:           1          4           SOC   1    0    0      2
5:           2          1         New 1   0    1    0      1
6:           2          2         New 1   0    1    0      1
   patient_wt age female grp_name
1:       0.25  55      1   Female
2:       0.25  65      1   Female
3:       0.25  55      0     Male
4:       0.25  65      0     Male
5:       0.25  55      1   Female
6:       0.25  65      1   Female
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> survmods_params <- create_params(psmfit_soc_wei, n = n_samples, 
+                                   uncertainty = "bootstrap", 
+                                   max_errors = 5, silent = TRUE)
R> survmods_params$pfs$coefs$scale <- cbind(survmods_params$pfs$coefs$scale,
+                                           coef_pfs_wei)
R> survmods_params$os$coefs$scale <- cbind(survmods_params$os$coefs$scale,
+                                          coef_os_wei)
R> survmods_params$pfs$coefs$scale[1:2, ]
\end{Sinput}
\begin{Soutput}
     (Intercept)     female      new1      new2
[1,]    1.784666 -0.1364781 0.1814189 0.2654431
[2,]    1.776743 -0.1173777 0.1390351 0.2907570
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> survmods_data[, ("(Intercept)") := 1]
R> survmods <- create_PsmCurves(survmods_params, 
+                               input_data = survmods_data)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> utilitymod <- create_StateVals(utility_tbl, n = n_samples, 
+                                 hesim_data = hesim_dat)
R> drugcostmod <- create_StateVals(drugcost_tbl, n = n_samples, 
+                                  hesim_data = hesim_dat)
R> medcostmod <- create_StateVals(medcost_tbl, n = n_samples, 
+                                 hesim_data = hesim_dat)
R> costmods <- list(Drug = drugcostmod, Medical = medcostmod)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> psm <- Psm$new(survival_models = survmods,
+                 utility_model = utilitymod,
+                 cost_models = costmods)
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> times <- seq(0, 30, by = .1)
R> psm$sim_survival(t = times)
R> head(psm$survival_)
\end{Sinput}
\begin{Soutput}
   sample strategy_id patient_id grp_id patient_wt curve   t
1:      1           1          1      1       0.25     1 0.0
2:      1           1          1      1       0.25     1 0.1
3:      1           1          1      1       0.25     1 0.2
4:      1           1          1      1       0.25     1 0.3
5:      1           1          1      1       0.25     1 0.4
6:      1           1          1      1       0.25     1 0.5
    survival
1: 1.0000000
2: 0.9998540
3: 0.9993126
4: 0.9982992
5: 0.9967670
6: 0.9946815
\end{Soutput}
\end{Schunk}

\begin{Schunk}
\begin{Sinput}
R> autoplot(psm$survival_, 
+           labels = c(labs_cohort,
+                      list(curve = c("PFS" = 1, "OS" = 2))
+                      ),
+           ci = TRUE, ci_style = "ribbon")
\end{Sinput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> psm$sim_stateprobs()
R> psm$stateprobs_[sample == 1 & patient_id == 1 & state_id == 2 & t == 12]
\end{Sinput}
\begin{Soutput}
   sample strategy_id patient_id grp_id patient_wt state_id  t
1:      1           1          1      1       0.25        2 12
2:      1           2          1      1       0.25        2 12
3:      1           3          1      1       0.25        2 12
        prob
1: 0.3215069
2: 0.3934037
3: 0.4148810
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
R> psm$sim_qalys(dr = .03) 
R> psm$sim_costs(dr = .03)
\end{Sinput}
\end{Schunk}
